\begin{abstract}
Any system that persists against entropy must minimize its rate of internal dissipation. We formalize this universal requirement as the Entropic Action ($S_\Phi = \int \mathcal{D} \, d\mu$), defined as the dissipation density $\mathcal{D} = \xi \cdot \mathbf{K} \cdot f$: the product of state transition cost ($\xi$), structural complexity ($\mathbf{K}$), and kinematic flux ($f$), integrated over the system's substrate. The Principle of Least Action emerges not as an irreducible axiom of mechanics, but as the necessary thermodynamic condition of minimum entropy production required for any system to resist dissolution.

While this principle applies universally, we validate it here against the most demanding test case: the quantum vacuum. Minimizing the Entropic Action on the discrete $E_8$ substrate (derived in \textit{Informational Energetics: A Universal Architecture of Persistence}) yields the unique Entropic Lagrangian, which strictly defines the functional operator forms corresponding to the Standard Model and General Relativity. The local hardware limits of the substrate (bandwidth and channel capacity) strictly dictate the functional form of each field operator: Capacity requires the fermion mass term; Map requires the Dirac operator for chiral propagation; Protocol necessitates Yang-Mills gauge fields to penalize phase decoherence; Governor demands the Higgs potential to enforce the finite state-space capacity; and Toll and Margin mandate the Einstein-Hilbert action as the bulk entropic cost of spacetime geometry. The finite capacity of the lattice rigorously bounds all operators to mass dimension 4, rendering the Entropic Lagrangian structurally closed. This closure prohibits higher-dimensional operators, higher-spin fields ($>2$), and dynamical dark energy by geometric necessity, not assumption. The Matter/Radiation distinction emerges as the two asymptotic minima of the dissipation functional. These structures are not postulated to fit empirical data, but derived as the unique, unavoidable architectural requirements for persistence.
\end{abstract}