\section{Introduction}

In \textit{Informational Energetics: A Universal Architecture of Persistence} \cite{meyer_informational_2026}, we introduced Informational Energetics (IE) and established that any persistent system, from a biological cell to the quantum vacuum, must resist entropic decay by satisfying six architectural pillars derived from control theory, information theory, and thermodynamics: Capacity (the vessel), Map (the model), Protocol (the interface), Governor (the constraint), Toll (the temporal cost), and Margin (the buffer). Applying these constraints to the vacuum uniquely determined the $E_8$ lattice as the geometric substrate of reality, validated by a parameter-free derivation of the fine-structure constant. This paper formalizes the dynamic Entropic Action and demonstrates that its minimization on this substrate uniquely yields the Standard Model Lagrangian and General Relativity.

\subsection{Entropic Action}
The central imperative of Informational Energetics is the \textbf{Selection Principle}: \textit{Among all possible configurations, only those that minimize Entropic Action ($S_\Phi$) relative to their structural complexity remain observable.} The Entropic Action $S_\Phi$ quantifies the total thermodynamic cost for a system to maintain structural coherence against environmental entropy over its persistence lifetime.

In \cref{sec:entropic_action}, we formalize the Selection Principle by defining the Entropic Action: $S_\Phi = \int \mathcal{D} \, d\mu$ where \textbf{Dissipation Density} $\mathcal{D} = \xi \cdot \mathbf{K} \cdot f$ is the product of the action cost per state transition ($\xi$), the structural complexity ($\mathbf{K}$), and the transition rate ($f$). We demonstrate that the Principle of Least Action is not an independent axiom of mechanics but a thermodynamic necessity: the strict condition of Minimum Entropy Production required for a system to survive.

\subsection{The Falsification Test: The Vacuum Substrate}
To validate this universal architecture, we apply it to the most demanding test case: the quantum vacuum. While open systems (such as biological organisms) can offset their action cost by maintaining a non-equilibrium steady state through a continuous flux of energy and matter, the vacuum possesses no external environment. To persist rather than dissolving into thermal noise, its optimization criterion is absolute: it must strictly minimize its internal rate of dissipation.

Applying this absolute thermodynamic constraint provides a structural solution to a longstanding conceptual puzzle: the origin of the Standard Model Lagrangian. While it is one of the most successful mathematical structures in the history of science, standard Quantum Field Theory (QFT) cannot explain \textit{why} nature chooses this specific form. It postulates the gauge symmetries, enforces renormalizability as a mathematical requirement rather than a physical limit, and inserts the parameter set $\mathcal{P} = \{g, \lambda, m, v\}$ entirely by hand.

\subsubsection{The \texorpdfstring{$E_8$}{E8}-Persistence Theory}
The $E_8$-Persistence Theory is the specific application of IE to fundamental physics\cite{meyer_informational_2026}. The following systems were derived, each building on the previous:

\begin{enumerate}
    \item \textbf{System 0: Requirements.} The six pillars were mapped to physics, deriving Finiteness, Unitarity, and Causality as existential necessities.
    \item \textbf{System 1: Substrate.} System 0's requirements uniquely specify an $E_8$ lattice projected to $D=4$ spacetime, yielding the characteristic integers $\mathbb{S} =$ $\{\Delta{=}43\text{ (temporal)},$ $\nu{=}16\text{ (chiral)},$ $\sigma{=}5\text{ (interaction)},$ $\chi{=}2\text{ (topological)}\}$.
    \item \textbf{System 2: Impedance.} The geometric impedance of this substrate yields the fine-structure constant $\alpha^{-1} = 137.035999212\dots$ to within $0.6\sigma$ of measurement, with zero free parameters.
\end{enumerate}
 
This paper derives System 3: the dynamic field equations. Because information is physical, every state transition carries an irreducible thermodynamic cost. The physical Principle of Least Action is identically the thermodynamic minimization of this information loss. If this substrate is the hardware of reality, minimizing Entropic Action across it must yield the known laws of physics. This paper performs that test. Because the six pillars encompass both internal field dynamics and spacetime geometry, the resulting \textbf{Entropic Lagrangian} unifies the Standard Model with General Relativity in a single variational principle.

\subsection{Structure of the Derivation}
The derivation proceeds by architectural necessity: Capacity requires mass; Map requires propagation; Protocol requires gauge coordination; Governor requires stability bounds; Toll and Margin require spacetime geometry. The resulting structure is closed; each pillar mandates a specific term in the Lagrangian, leaving no freedom for arbitrary additions or omissions.

\Cref{sec:entropic_lagrangian} maps the Entropic Action to the $E_8$ substrate, formulating the covariant action $S_\Phi = \int d^4x \sqrt{-g} \mathcal{L}_\Phi$, defining kinematic and thermodynamic bounds, and translating  discrete information-theoretic costs into continuous field operators. \Cref{sec:map,sec:protocol,sec:governor,sec:toll_margin} then systematically derive the field equations for each architectural pillar.

This structural derivation carries profound conceptual consequences absent from standard QFT: a physical origin for the Matter/Force distinction, a structural explanation for UV completion, and rigid, falsifiable predictions (\cref{sec:falsifiability}).

\textit{Scope and Dependencies:} To maintain rigorous boundaries and keep the mathematical derivations focused, this paper categorizes three distinct domains:
\begin{enumerate}
    \item \textbf{Established (The Substrate):} The $E_8 \to D_4 \oplus D_4$ geometric substrate and its characteristic invariants $\mathbb{S}=\{\Delta=43,\nu=16,\sigma=5,\chi=2\}$ are treated as proven axioms derived from Informational Energetics\cite{meyer_informational_2026}.
    \item \textbf{Derived Here (The Functional Forms):} This paper strictly derives the \textit{structural forms} of the physical operators (the Entropic Lagrangian). We demonstrate that the architectural limits of the substrate uniquely mandate the Dirac, Yang-Mills, Higgs, and Einstein-Hilbert actions.
    \item \textbf{Future Work (The Numerical Constants):} Standard QFT requires the manual insertion of continuous parameters (e.g., the specific running gauge couplings $g_1, g_2, g_3$, the Higgs parameters $\mu$ and $\lambda$, and the Planck Mass $M_P$ and $\Lambda$ (cosmological constant)). Because this framework derives physics from a discrete geometry, these values are not free parameters. However, their explicit derivation from the substrate integers is deferred entirely to the companion paper, \textit{The Geometric Constants} (in progress).
\end{enumerate}