\section{Derivation D: Spacetime Dynamics (Toll \& Margin)}
\label{sec:spactime}

The final sector of the Lagrangian concerns the substrate itself. Unlike standard Quantum Field Theory, which assumes a fixed background, the Entropic Lattice is a dynamic entity. It must expend action to maintain the manifold geometry ($D=4$) against the entropic tendency to decohere into a random graph.

This structural maintenance corresponds to the final two pillars: \textbf{Toll} (The cost of transitions) and \textbf{Margin} (The cost of existence).

\subsection{The Toll: The Ricci Curvature (\texorpdfstring{$R$}{R})}
The \textbf{Toll} pillar represents the irreversible entropic cost of state transitions (Time). In a geometric theory, the flow of time is defined by the metric tensor $g_{\mu\nu}$.

If the lattice is perfectly flat (Minkowski), the update cycles are uniform. However, the presence of information (matter/energy) creates a \textbf{Capacity Deficit} (derived in System VI of the companion paper). To maintain the constant speed of information transfer ($c=1$), the lattice geometry must deform.

The entropic cost of this deformation is measured by the \textbf{Riemann Curvature}. The simplest scalar invariant representing the "bending" of the update trajectory is the Ricci Scalar $R$. The action cost is proportional to the total curvature over the manifold volume:
\begin{equation}
    \mathcal{L}_{Toll} = \frac{M_P^2}{2} R
\end{equation}
where $M_P$ (The Planck Mass) acts as the \textbf{Bulk Modulus} or stiffness of the lattice, converting geometric curvature ($L^{-2}$) into energy density ($M^4$).

\subsection{The Margin: The Cosmological Constant (\texorpdfstring{$\Lambda$}{Lambda})}
The \textbf{Margin} pillar represents the resolution floor. A system with zero signal margin is indistinguishable from noise.

The vacuum is not a continuum; it is a discrete lattice with finite node spacing. Consequently, the "empty" state is not zero energy, but the \textbf{Minimum Resolvable Energy} required to define a bit of spatial volume.
This manifests as a constant energy density penalty for maintaining the volume of the universe. In General Relativity, this is the Cosmological Constant $\Lambda$:
\begin{equation}
    \mathcal{L}_{Margin} = - \frac{M_P^2}{2} (2\Lambda) = - \rho_{vac}
\end{equation}
The negative sign indicates that this is a cost paid \textit{per unit volume}. The vacuum must "pay rent" in entropic action to maintain a coherent spatial manifold.

\subsection{Synthesis: The Einstein-Hilbert Action}
Combining the Toll (Curvature) and Margin (Volume Cost), we recover the Einstein-Hilbert action as the unique solution for the bulk regulator:
\begin{equation}
    S_{Gravity} = \int d^4x \sqrt{-g} \frac{M_P^2}{2} (R - 2\Lambda)
\end{equation}
\textbf{Conclusion:} General Relativity is not a separate force. It is the \textbf{Operating System} of the lattice. The Einstein-Hilbert action describes the metabolic cost of maintaining the spacetime container (Margin) and updating the metric (Toll) to preserve causal order.