\section{Falsifiability: Structural Kill-Switches}
\label{sec:falsifiability}

Unlike Effective Field Theories, which can accommodate new phenomena by adding arbitrary higher-dimensional operators and continuous tuning parameters, the Entropic Lagrangian is structurally closed. The Characteristic Integers $\mathbb{S} = \{\Delta, \nu, \sigma, \chi\}$ completely saturate the geometry's degrees of freedom. Therefore, the theory presents specific ``Hard Constraints'' that serve as immediate falsification criteria. 

The following criteria test distinct aspects of the framework. Criteria 1, 2, 4, and 6 test the \textit{functional forms} derived entirely within this paper. Criteria 3 and 5 test the \textit{specific integer values} and structural invariants of the $E_8$ substrate:

\begin{enumerate}    
    \item \textbf{Lorentz Violation at Low Energy:} The derivation of the Metric Signature relies on the Bandwidth Limit ($c_{eff}=1$). Any experimental confirmation of Lorentz invariance violation (e.g., energy-dependent speed of light) at energies $E \ll M_P$ would invalidate the channel capacity argument.

    \item \textbf{The Prohibition of Other Spins:} The substrate geometry strictly prohibits fundamental particles with $spin > 2$ (and fundamental half-integer $spin > 1/2$), providing a structural mechanism for the Weinberg-Witten theorem:
    \begin{itemize}
        \item \textbf{No Supersymmetry / Supergravity (Spin-3/2):} A fundamental spin-3/2 field (gravitino) requires a vector-spinor coupling. As established in the substrate projection, this cannot be constructed from the orthogonal $D_4 \oplus D_4$ decomposition without introducing geometrically suppressed mirror fermions. This structurally falsifies Supersymmetry as a fundamental property of the vacuum.
        \item \textbf{No Spin $\ge$ 3:} A fundamental spin-3 field in 4D spacetime requires a totally symmetric rank-3 tensor. Off-shell, before classical trace constraints are applied, this unconstrained tensor possesses exactly 20 independent components. Because the fundamental chiral bit-depth of the lattice node is strictly limited to $\nu=16$, the substrate lacks the hardware channel capacity to encode the off-shell quantum fluctuations of a fundamental spin-3 state. Higher-spin extensions are therefore physically forbidden by Shannon capacity limits.
    \end{itemize}
    
    \item \textbf{The Fourth Generation Prohibition:} A single Standard Model generation (including right-handed neutrinos) requires exactly 16 chiral Weyl fermions. The internal chiral channel capacity is exactly saturated by this payload ($\nu=16$). The replication of this structure into exactly 3 generations is geometrically fixed by the unique Triality of the internal $D_4$ sector (representing three distinct outer-automorphism framings of the same 16-channel hardware). The discovery of a 4th generation of fermions would require breaking this Triality or implying $\nu > 16$, either of which strictly falsifies the $E_8 \to D_4 \oplus D_4$ substrate hypothesis.
    
    \item \textbf{The Perturbative Gravity Prohibition:} We derive General Relativity as an emergent hydrodynamic limit (\cref{sec:toll_margin}). This requires that gravity be asymptotically safe or non-perturbative in the UV. The discovery that gravity follows standard perturbative quantization (like QED) at the Planck scale would falsify the bulk regulator derivation.

    \item \textbf{The Lattice Strain Check ($\kappa \neq 1$):} The derivation of Gravity as an emergent compensating strain field closes the Einstein-Hilbert limit algebraically, but this requires non-perturbative numerical confirmation. A full non-perturbative Monte Carlo simulation of the capacity-conserving $E_8$ substrate must reproduce a massless spin-2 correlator with the exact Einstein-Hilbert normalization. If lattice artifacts introduce a mass gap for the graviton or disrupt the macroscopic strain response to local capacity deficits, yielding a dimensionless stiffness ratio $\kappa \neq 1$ (outside finite-size scaling errors), it would explicitly falsify the emergence of General Relativity from this specific discrete substrate.

    \item \textbf{Dynamical Dark Energy ($\Lambda$ Evolution):} The Margin pillar strictly defines the Cosmological Constant as a static, non-derivative geometric baseline cost required to resolve the spatial manifold. Any confirmed cosmological detection of an equation of state parameter $w \neq -1$, or a dynamically evolving dark energy component ($w_a \neq 0$), would falsify the identification of $\Lambda$ as the static Margin cost.
\end{enumerate}